---
title: "Project 1 Reproducible Research"
author: "Illya Pavao Bjazevic"
date: "9/8/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("tidyverse")
```

### 1. Code for reading in the dataset and/or processing the data.

We need to download all the data that we will user for this homework.



```{r}
# Download the data from the url of the homework.

download.file("https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip", destfile = "myData.zip", mode="wb")

# Unzip the file, 
# if we have not already extracted the CSV file 
# activity.csv
# Otherwise, do nothing.

if (!file.exists("activity.csv") ) {unzip("myData.zip")}

originalDataFrame <- read.csv("activity.csv", header = TRUE)


#We create a DataFrame without NA values

originalDataframeWITHOUTnaVALUES <- originalDataFrame[!is.na(originalDataFrame$steps),]

# We create a DATAFRAME with the steps per day
# using the 
# Aggregate function

TOTALstepsPerDayDATAFRAMEwithoutNAvalues <- aggregate(steps ~ date, originalDataframeWITHOUTnaVALUES, FUN=sum)

```

### This is the data that we analyze in this project. This is a DATAFRAME without the NA values that contains a column with the total of STEPS PER DAY.

```{r}
#1. This is the data that we analyze in this project. This is a DATAFRAME without the NA values that contains a column with the total of STEPS PER DAY.

str(TOTALstepsPerDayDATAFRAMEwithoutNAvalues)
head(TOTALstepsPerDayDATAFRAMEwithoutNAvalues)
```

### 2.Histogram of the total number of steps taken each day of TOTALstepsPerDayDATAFRAMEwithoutNAvalues.

```{r}
#2. Histogram of the total number of steps taken each day of TOTALstepsPerDayDATAFRAMEwithoutNAvalues.

hist(TOTALstepsPerDayDATAFRAMEwithoutNAvalues$steps, main = "Histogram TOTALstepsPerDayDATAFRAMEwithoutNAvalues", xlab = "TOTALstepsPerDay")
```

### 3.Mean and median TOTALstepsPerDayDATAFRAMEwithoutNAvalues.

```{r}
#3. Mean and median TOTALstepsPerDayDATAFRAMEwithoutNAvalues.

mean(TOTALstepsPerDayDATAFRAMEwithoutNAvalues$steps)
median(TOTALstepsPerDayDATAFRAMEwithoutNAvalues$steps)

```

### 4.Time series plot of the average number of steps taken originalDataframeWITHOUTnaVALUES

```{r}
# 4.Time series plot of the average number of steps taken originalDataframeWITHOUTnaVALUES

averageTOTALstepsPerDayDATAFRAMEwithoutNAvalues <- aggregate(steps ~ interval, originalDataframeWITHOUTnaVALUES, mean)

str(averageTOTALstepsPerDayDATAFRAMEwithoutNAvalues)
head(averageTOTALstepsPerDayDATAFRAMEwithoutNAvalues)

plot(averageTOTALstepsPerDayDATAFRAMEwithoutNAvalues$interval, averageTOTALstepsPerDayDATAFRAMEwithoutNAvalues$steps, type='l', col=1, main="averageTOTALstepsPerDayDATAFRAMEwithoutNAvalues")


```

### 5.The 5-minute interval that, on average, contains the maximum number of steps

```{r}

#5.The 5-minute interval that, on average, contains the maximum number of steps. (MAXIMUM VALUE AND POSITION OF THE MAXIMUM VALUE)

#MAXIMUM VALUE
MAXIMUM <-max(averageTOTALstepsPerDayDATAFRAMEwithoutNAvalues$steps)
MAXIMUM

#POSITION OF THE MAXIMUM VALUE
which(grepl(MAXIMUM, averageTOTALstepsPerDayDATAFRAMEwithoutNAvalues$steps))

```

### 6.Replace NA values with the mean of the set. Code to describe and show a strategy for imputing missing data



```{r}

#6.Replace NA values with the mean of the set. Code to describe and show a strategy for imputing missing data

MEANofTHEset<-mean(originalDataframeWITHOUTnaVALUES$steps)
MEANofTHEset

#x[is.na(x)] <- 0
OPTIMIZATIONoriginalDataFrame <- read.csv("activity.csv", header = TRUE)
OPTIMIZATIONoriginalDataFrame$steps[is.na(OPTIMIZATIONoriginalDataFrame$steps)]<-MEANofTHEset
str(OPTIMIZATIONoriginalDataFrame)
head(OPTIMIZATIONoriginalDataFrame)

```


### 7.Histogram of the total number of steps taken each day after missing values are imputed

```{r}
#7.Histogram of the total number of steps taken each day after missing values are imputed

hist(OPTIMIZATIONoriginalDataFrame$steps, main = "Histogram OPTIMIZATIONoriginalDataFrame", xlab = "TOTALstepsPerDay")

```

### 8.Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends

```{r}
#8.Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends
library(chron)
#x[is.weekend(x)] check if a day is a weekend
OPTIMIZATIONoriginalDataFrame$date[is.weekend(OPTIMIZATIONoriginalDataFrame$date)]<- 'Weekend'
OPTIMIZATIONoriginalDataFrame$date[!is.weekend(OPTIMIZATIONoriginalDataFrame$date)]<- 'Weekday'

weekendDATAFRAME<-OPTIMIZATIONoriginalDataFrame[OPTIMIZATIONoriginalDataFrame$date == 'Weekend',]

averageWEEKEND <- aggregate(steps ~ interval, weekendDATAFRAME, mean)

plot(averageWEEKEND$interval, averageWEEKEND$steps, type='l', col=1, main="averageWEEKEND")

weekdayDATAFRAME<-OPTIMIZATIONoriginalDataFrame[OPTIMIZATIONoriginalDataFrame$date == 'Weekday',]

averageWEEKDAY <- aggregate(steps ~ interval, weekdayDATAFRAME, mean)

plot(averageWEEKDAY$interval, averageWEEKDAY$steps, type='l', col=1, main="averageWEEKDAY")


```

